---
title: "Linear Regression"
author: "Baptiste CRINIERE-BOIZET"
institute: Data Analysis Core
output: binb::metropolis
    # beamer_presentation:
    # theme: "metropolis"
header-includes:
  - \definecolor{mycolor}{HTML}{e23d18}
  - \definecolor{mycolor2}{HTML}{002e82}
  - \setbeamercolor{title}{fg=mycolor2}
  - \setbeamercolor{frametitle}{fg=white}
  - \setbeamercolor{progress bar}{fg=mycolor}
  - \setbeamercolor{block title}{fg=mycolor}
  - \setbeamercolor{normal text}{fg=mycolor2}
  - \setbeamercolor{palette primary}{fg=white, bg=mycolor2}
  - \setbeamercolor{palette secondary}{fg=mycolor, bg=mycolor}
  - \setbeamercolor{itemize item}{fg=mycolor}
  - \setbeamercolor{itemize subitem}{fg=mycolor}
  - \setbeamercolor{itemize subsubitem}{fg=mycolor}
  - \setbeamercolor{enumerate item}{fg=mycolor}
  - \setbeamercolor{enumerate subitem}{fg=mycolor}
  - \setbeamercolor{enumerate subsubitem}{fg=mycolor}
  - \setbeamercolor{footline}{fg=mycolor}
  - \titlegraphic{\vspace{7cm}\flushright\includegraphics[width=4cm]{ICM_rvb.png}}
  # - \usepackage{animate}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.align = "center")
knitr::opts_chunk$set(out.width="110%")
```

\tableofcontents[sectionstyle=show, subsectionstyle=hide]

# Theory
```{r, results='hide'}
library(tidyverse)
library(ggpubr)
library(MASS)
library(ggsci)
library(emmeans)
library(plotly)
set.seed(1234)
X <- rnorm(100, mean = 175, sd = 10)
Res <- rnorm(100, mean = 0, sd = 5)
Y <- X + Res
data <- data.frame(X, Y)
mod <- lm(Y ~ X, data)
summary(mod)
```


## Dependent Variable and Independent Variables



- Key Distinction: Dependent Variable (Y) vs Independent Variables (X1, X2, etc.)
\vspace{0.5cm}
- Dependent Variable = Response or Explained Variable
\vspace{0.5cm}
- Independent Variable = Predictor or Explanatory
\vspace{0.5cm}
- Choice based on logic or theory, for example: child's height depending on the mother's height

